Notes and other text

### 2023-03-31 doc/manuscript/translocation.qmd
Explore structures for opening paragraphs in discussion

While conclusively demonstrating that evolutionary rescue is driving population recoveries is challenging.
we provide four lines of evidence that evolutionary rescue is a strong candidate mechanism for observed resistance and recoveries in MYL frog systems.
First, in host-pathogen systems, evolutionary rescue is more likely to occur when hosts experience repeated selection events (e.g., through re-infections), there is a persistent pathogen reservoir or frequency-dependent infection dynamics, and populations experience increased per capita birth rates at small population sizes [@golas2021].
At least two of these conditions are met in the MYL frog system: i) individuals that recover from Bd infection do not gain long-lasting immunity to Bd and are susceptible to re-infection and ii) there are potential Bd reservoirs (Briggs, Reeder) and seasonal frequency-dependent transmission dynamics (Wilber) in the MYL frog system.
Thus, MYL frog-Bd systems have the epidemiological characteristics necessary for evolutionary rescue to occur.
Second, our genomic analyses comparing naive and recovering MYL frog populations identified notable differences in allele frequencies between adjacent Bd-naive and recovering population.
These differences are consistent with strong signatures of selection and regions of the genome under selection are associated with known mechanisms of resistance to Bd infection.
Third, our translocations allowed us to conclude that observed resistance and recovery of resistant populations is not strictly due to local environmental conditions (though environment does play a role).
Changes in population growth rate due to evolved resistance extend beyond the natal environment and into subsequent generations, as demonstrated by high survival rates of translocated frog populations in new habitats in the presence of Bd.
Finally, evolved resistance has a quantifiable phenotype.
In persistent populations and those established via translocation of frogs from recovering populations, Bd load is not an important predictor of post-translocation frog survival (in contrast to important relationship between Bd load and frog survival during Bd-caused epizootics).
While other studies have documented signatures of selection in hosts following disease-induced declines (Epstein, white-nose, Byrne), our study is unique in that we show that evolved resistance is associated with population recovery in both natal and non-natal habitats, providing an actionable strategy for promoting the long-term viability of a species.

Despite the results we present here, recovery of Bd-impacted amphibians worldwide is expected to be challenging at best.
Hundreds of declining species, but mostly failed attempts at recovery.
Failures may be due in part to a general inability by amphibians to mount an effective immune response against Bd infection.
In contrast, in our study MYL frogs (a species highly susceptible to chytridiomycosis) showed relatively high frog survival despite ongoing Bd infection and low probability of extinction for many reestablished populations.

>>>>

While evolutionary rescue is not the only mechanism that can lead to host recovery in the presence of disease (Brannely), there are four lines of evidence that evolutionary rescue is a strong candidate mechanism for observed resistance and recoveries in MYL frog systems.
First, in host-pathogen systems, evolutionary rescue is more likely to occur when hosts experience repeated selection events (e.g., through re-infections), there is a persistent pathogen reservoir or frequency-dependent infection dynamics, and populations experience increased per capita birth rates at small population sizes [@golas2021].
At least two of these conditions are met in the MYL frog system: i) individuals that recover from Bd infection do not gain long-lasting immunity to Bd and are susceptible to re-infection and ii) there are potential Bd reservoirs (Briggs, Reeder) and seasonal frequency-dependent transmission dynamics (Wilber) in the MYL frog system.
Thus, MYL frog-Bd systems have the epidemiological characteristics necessary for evolutionary rescue to occur.
Second, our genomic analyses comparing naive and recovering MYL frog populations identified notable differences in allele frequencies between adjacent Bd-naive and recovering population.
These differences are consistent with strong signatures of selection and regions of the genome under selection are associated with known mechanisms of resistance to Bd infection.
Third, our translocations allowed us to conclude that observed resistance and recovery of resistant populations is not strictly due to local environmental conditions (though environment does play a role).
Changes in population growth rate due to evolved resistance extend beyond the natal environment and into subsequent generations, as demonstrated by high survival rates of translocated frog populations in new habitats in the presence of Bd.
Finally, evolved resistance has a quantifiable phenotype.
In persistent populations and those established via translocation of frogs from recovering populations, Bd load is not an important predictor of post-translocation frog survival (in contrast to important relationship between Bd load and frog survival during Bd-caused epizootics).
While other studies have documented signatures of selection in hosts following disease-induced declines (Epstein, white-nose, Byrne), our study is unique in that we show that evolved resistance is associated with population recovery in both natal and non-natal habitats, providing an actionable strategy for promoting the long-term viability of a species.

Despite the results we present here, recovery of Bd-impacted amphibians worldwide is expected to be challenging at best.

-   Hundreds of declining species, but mostly failed attempts at recovery.
    Failures may be due in part to a general inability by amphibians to mount an effective immune response against Bd infection.
    In contrast, in our study MYL frogs (a species highly susceptible to chytridiomycosis) showed relatively high frog survival despite ongoing Bd infection and low probability of extinction for many reestablished populations.
